---
title: "p8105_hw2_yz4437"
author: "Yuchen Zhang"
date: "2022-10-04"
output: github_document
---

```{r load_libraries}
library(tidyverse)
library(readxl)
```

### Problem 1


Read and clean the data.he process begins with data import, updates variable names, and selects the columns.update `entry` from `yes` / `no` to a logical variable. 
```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(
    entry = ifelse(entry == "YES", TRUE, FALSE)
    )
```
these data are not "tidy":  we need to convert `route` variables from wide to long format. The dimension of the resulting dataset is `r nrow(trans_ent)` x `r ncol(trans_ent)`.

```{r}
dist_station = distinct(trans_ent,line, station_name,.keep_all = T)
```
There are `r nrow(dist_station)` distinct stations.  
`r nrow(filter(dist_station,ada==T))` stations are ADA compliant. 
`r round(nrow(filter(trans_ent,vending == "NO", entry == T))/nrow(trans_ent),2) station entrances / exits without vending allow entrance.

```{r}
trans_ent_ref = 
  trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct
```
There are `r nrow(trans_ent)` stations serving the A train, and `r nrow(filter(trans_ent, ada==T))` of them are ADA compliant.


### Problem 2

load data from `Mr. Trash Wheel` sheet from file `Trash-Wheel-Collection-Totals-7-2020-2.xlsx`.  omit non-data entries by range, update variable names, omit rows that do not include dumpster-specific data, round the number of sports balls to the nearest integer and converts the result to an integer variable.
```{r problem2}

Mr.trash_df = read_excel("data/Trash Wheel Collection Data.xlsx",
      sheet = "Mr. Trash Wheel",
      range = "A2:N549") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))

Mr.trash_df

```

similar process to `Professor Trash Wheel`
```{r}

ProTrash_df = read_excel("data/Trash Wheel Collection Data.xlsx",
      sheet = "Professor Trash Wheel", 
      range = "A2:M96") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster)

ProTrash_df
```
add an additional variable to both datasets `trash_wheel_name` to keep track which trash wheel is and combine these two datasets. column type of `years` in `Mr.trash_df` is changed from character to double for same type in both datasets. 

```{r}

Mr.trash_df = mutate(Mr.trash_df, trash_wheel_name = "mr.") %>% 
    mutate(
    year = as.double(year)
    ) 

ProTrash_df = mutate(ProTrash_df, trash_wheel_name = "professor") 

total_df = full_join(Mr.trash_df, ProTrash_df)
```
the number of observations in the resulting dataset is `r nrow(total_df)`, have key variables Trash Wheel name, dumpster number, date of trash collection, weight and volume , and 8 types of trash. 

The total weight of trash collected by Professor Trash Wheel is `r sum(filter(total_df, trash_wheel_name == "professor")$weight_tons)` tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(filter(total_df, trash_wheel_name == "mr.", year == 2020)$sports_balls)`.


### Problem 3

separate variable `mon` into integer variables `year`, `month` and `day`. convert `month` type from number to month abbreviation. create a `president` variable taking values `gop` and `dem`, and remove `prez_dem`, `prez_gop` and `day` variables.
```{r problem3_step1}

pols = read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>%
    janitor::clean_names() %>%
    separate(mon,c('year', 'month','day'), sep = '-') %>%
    mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day),
    month = month.abb[month],
    president = ifelse(prez_dem == 1,"dem","gop")
    ) %>% 
    select(-prez_dem, -prez_gop, -day)

pols

```

separate variable `date` into integer variables `month`, `day`and `year`. arrange according to `year` and `month`, and move `year` and `month` to leading columns.

```{r problem3_step2}

snp = read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>%
    janitor::clean_names() %>%
    separate(date, c('month', 'day','year'), sep = '/') %>%
    mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day),
    month = month.abb[month],
    year = ifelse(year < 22,year + 2000,year + 1900)
    ) %>% 
    arrange(year,month) %>% 
    select(year,month,close)
snp

```
switch dataset from wide to long format. create variables `month`, values are month abbreviation from original column names and create variable `percent_of_unemployment` values are from sheet.
```{r problem3_step3}
  unemp = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
    pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "percent_of_unemployment"
    )%>%
    janitor::clean_names() %>% 
    mutate(year = as.integer(year)) 
```

merging snp into pols, and merging unemployment into the result.
```{r}
snp_pols = left_join(pols,snp)
result = left_join(snp_pols,unemp)
```

Dataset `pols` has dimension `r nrow(pols)` with `r ncol(pols)`, with year range from `r min(pols$year)` to `r max(pols$year)`. It contains the number of republican/democratic in governors, senators, representatives and president on the associated date.

Dataset `snp` has dimension `r nrow(snp)` with `r ncol(snp)`, with year range from `r min(snp$year)` to `r max(snp$year)`. It contains the closing values of the S&P stock index on the associated date.

Dataset `unemp` has dimension `r nrow(unemp)` with `r ncol(unemp)`, with year range from `r min(unemp$year)` to `r max(unemp$year)`. It contains percentage of unemployment in that associated year and month.

Dataset `result` has dimension `r nrow(result)` with `r ncol(result)`, with year range from `r min(result$year)` to `r max(result$year)`. It contains number of certain party in government, S&P stock index and unemployment rate that associated year and month in one dataset.
